---
title: "Predict wine quality"
author: "Peter Strauch"
date: "15. októbra 2018"
output:
  html_document:
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = TRUE, fig.align='center')
```

### References
How to do a reproducible document:  
[>  Markdown Basics](https://rmarkdown.rstudio.com/authoring_basics.html)  
[>  R Reference Card](https://cran.r-project.org/doc/contrib/Short-refcard.pdf)  
[>  The caret Package](https://topepo.github.io/caret/index.html)

How to obtain data:  
[> 24 Ultimate Data Science Projects To Boost Your Knowledge and Skills](https://www.analyticsvidhya.com/blog/2018/05/24-ultimate-data-science-projects-to-boost-your-knowledge-and-skills/)  
[>  The UCI Machine Learning Repository](http://archive.ics.uci.edu/ml/datasets.html)  


### Load data and libraries
[About data](http://archive.ics.uci.edu/ml/datasets/Wine+Quality)
```{r loadings}
library(caret)
library(rattle)
library(ggplot2)
library(randomForest)
library(rpart)
library(gbm)
library(nnet)

## read data from URLs
url1 <- "http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv"
url2 <- "http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv"
data_red <- read.csv(url(url1), sep = ";")
data_white <- read.csv(url(url2), sep = ";")

## dimensions of data
dim(data_red)
dim(data_white)
```

Input variables (based on physicochemical tests):  

+ 1. fixed acidity  
+ 2. volatile acidity  
+ 3. citric acid  
+ 4. residual sugar  
+ 5. chlorides  
+ 6. free sulfur dioxide  
+ 7. total sulfur dioxide  
+ 8. density  
+ 9. pH  
+ 10. sulphates  
+ 11. alcohol  

Output variable (based on sensory data):  

+ 12. quality (score between 0 and 10)


# Research question

+ Q: From what parameters depends the quality of red wine?
+ Classification problem: quality is categorical (factor, mark)
```{r data modification}
## only red wine
data <- data_red

## convert predicted outcome into factor variable
data$quality <- as.factor(data$quality)
```

+ I picked data about red wine because I like red wine more. :) 
+ If you are interested in white whine, use `data <- data_white`. 
+ It is also possible to mix data together by using `data <- rbind(data_red, data_white)` but this depend on data and research question.

Then, Approach is almost the same.


# Key idea of each ML algorithm

+ first, we split data into training and testing dataset,
+ then, we use training data for exploration and train a model
+ at the end, we evaluate model on testing dataset

```{r splitting data}
## do a split
set.seed(1111)
inTrain <- createDataPartition(data$quality, p = 0.7, list = F)
training <- data[inTrain,]
testing <- data[-inTrain,]

## check dimensions
dim(training)
dim(testing)
```
*Note:* I use `set.seed()` because of reproducibility - same results, but still random.


## Explore data
```{r data exploration, fig.width = 10, fig.height=8}
## info about data
str(training)
summary(training)

## plot matrix
plot(training[,-12], col = training$quality)
```

```{r data exploration-2}
## plot - alcohol vs. volatile.acidity
qplot(alcohol, volatile.acidity, data = training, col = quality, alpha=I(0.5))
```

We have too many categories for quality variable, so let's look only for data with `quality = 5 or 6 or 7` because here is the most of information.
```{r data exploration on reduced data, fig.width = 10, fig.height=8}
## reduced data
data_reduced <- training[training$quality==5 | training$quality==6 | training$quality==7,]

## plot matrix
plot(data_reduced[,-12], col = data_reduced$quality)
```

```{r data exploration on reduced data-2}
## plot - alcohol vs. volatile.acidity
g1 <- qplot(alcohol, volatile.acidity, data = data_reduced, col = quality, alpha=I(0.5))
g1
```

We can check, if some variable has near zero variance. If yes, we can omit this variable from the model building.  
We can also check, if some variable is corelated with other, and thus can be also omited.
```{r nzv}
## near zero variance
nzv(training, saveMetrics = TRUE)

## corelations
M <- cor(training[,-12])
M <- round(abs(M), digits = 2)
diag(M) <- 0
max(M)
```





## Prediction

### Guessing of some "areas" in graph with dominant quality  
Some magic :)
```{r guess a model}
g1 + geom_vline(xintercept = 10.25) 
```


## 1. Decision trees
first of all, we will try a tree - easy and interpretable model. We use `rpart` library.
```{r tree-build}
## build model - tree
model_tree <- train(quality ~., data = training, method="rpart")
```

Here is information about model. Accuracy of this model on testing data is **`r round(max(model_tree$results$Accuracy), digits=4)`**.
```{r tree-info}
## model info
model_tree

## plot - accuracy from tuning parameter (cp)
plot(model_tree)

## importance of variables in the model
varImp(model_tree)
plot(varImp(model_tree))
```

Decisions from the tree are:
```{r tree-decisions}
## tree
model_tree$finalModel
fancyRpartPlot(model_tree$finalModel)

## percentages in the tree
table(training$quality) / nrow(training)
```

But this model can be overfitted on training dataset (building the model on training data), so we need to chceck for accuracy in testing dataset (independent data from the model).
```{r tree-confusion matrix}
## predictions of quality on testing data
pred_tree <- predict(model_tree, testing)

## compare real quality with predicted values
confusionMatrix(testing$quality, pred_tree)
```

Accuracy on testing data is unbalanced with accuracy of the model (builded on training data).



### 1a. Decision tree with Cross-validation
Now, we use 7-fold cross-validation for building a model. The advantage is reduction of overfitting.
```{r tree with cv7}
set.seed(1234)
model_tree_cv7 <- train(quality ~., data = training, method="rpart", 
                        trControl = trainControl("cv", number = 7))
model_tree_cv7
confusionMatrix(testing$quality, predict(model_tree_cv7, testing))
fancyRpartPlot(model_tree_cv7$finalModel)
```

We also can check accuracy in each fold:
```{r tree with cv7-acc of folds}
model_tree_cv7$resample
```


### 1b. Decision tree with preprocessing of data
We can also preprocess data before building any model. We saw some skewness of data in exploration, so we can scale data.
```{r tree with preprocess}
set.seed(1234)
model_tree_scaled <- train(quality ~., data = training, method="rpart", 
                           trControl = trainControl("cv", number = 7),
                           preProcess = "scale")
model_tree_scaled
confusionMatrix(testing$quality, predict(model_tree_scaled, testing))[3]
```
Scaling and centring have no impact on a model.
*Note:* We can use wide variety of preprocessing, e.g. `PCA, scaling, centring, knnImpute, BoxCox,` ...


## 2. Improvement of trees
### 2a. Boosted with trees
Gradient boosting method. Idea of boosting is to do a linear combination of predictors.
```{r boosted tree-model, results='hide'}
## build model - boosted
set.seed(1234)
model_gbm <- train(quality ~., data = training, method="gbm", 
                  trControl = trainControl("cv", number = 7))
```

```{r boosted tree-info}
## info about model
model_gbm
plot(model_gbm)
summary(model_gbm)

## acc on testing data
confusionMatrix(testing$quality, predict(model_gbm, testing))
```

### 2b. Bagged trees
Idea of bagging is bootstrap - calculate model on resamples and take the average model (same bias, less variability).
```{r bagged tree}
set.seed(1234)
model_treebag <- train(quality ~., data = training, method="treebag", 
                       trControl = trainControl("cv", number = 7))

## info about model
model_treebag
model_treebag$finalModel

## acc on testing data
confusionMatrix(testing$quality, predict(model_treebag, testing))
```


## 3. Random forest
```{r rf-build}
## build model - random forest
set.seed(1234)
model_rf <- train(quality ~., data = training, method="rf", 
                  trControl = trainControl("cv", number = 7))

## info about model
model_rf
plot(model_rf)
varImp(model_rf)

## acc on testing data
confusionMatrix(testing$quality, predict(model_rf, testing))
```

We can check of how many trees are in the forest or check some tree.
```{r rf-trees}
## number of trees in rf
model_rf$finalModel

## 155th tree in rf
head(x = getTree(model_rf$finalModel, k=155), n = 15)
```


Here is how the error rate depends on number of trees. 
```{r rf-error}
## plot - error
plot(model_rf$finalModel)
```
We can see that is not neccessary to have so many trees in forest. Next time we can specify argument `ntrees=100` into train function and save time needed for model building.



## 4. Comparison with Neural networks
We can try to use neural network (which is "black box") to save the day.
```{r neural network-build, results='hide'}
## build model - neural network
set.seed(1234)
model_nnet <- train(quality ~., data = training, method="nnet", 
                    trControl = trainControl("cv", number = 7))
```

```{r neural network-info, echo=FALSE}
## info about model
model_nnet
plot(model_nnet)

## acc on testing data
confusionMatrix(testing$quality, predict(model_nnet, testing))
```
Neural networks looks cool and complex, nobody understand it ... but some(many-)times results are worse.


# Summary
More info about methods and more methods:
[The caret Package, cap.7, train Models By Tag](https://topepo.github.io/caret/train-models-by-tag.html)
```{r compare acc}
## simple tree
confusionMatrix(testing$quality, predict(model_tree, testing))$overall[1]

## cross-validated tree
confusionMatrix(testing$quality, predict(model_tree_cv7, testing))$overall[1]

## scaled tree
confusionMatrix(testing$quality, predict(model_tree_scaled, testing))$overall[1]

## boosted tree
confusionMatrix(testing$quality, predict(model_gbm, testing))$overall[1]

## bagged tree
confusionMatrix(testing$quality, predict(model_treebag, testing))$overall[1]

## random forest
confusionMatrix(testing$quality, predict(model_rf, testing))$overall[1]

## neural network
confusionMatrix(testing$quality, predict(model_nnet, testing))$overall[1]
```

And now, let's **predict quality** of "our" **new wine**:
```{r}
## our wine parameters (average value of each column)
our_wine <- as.data.frame(t(apply(X = data[,-12], MARGIN = 2, FUN = mean)))

## predict quality of a wine, which isn't asses yet
predict(object = model_rf, newdata = our_wine)
```




------
# Addition
2 more datasets from References at the top of this document:

+ Iris data - realy clear model
+ Loan prediction data - interesting


### Iris data
```{r iris}
## data
data("iris")

## exploratory
plot(iris, col=iris$Species)
qplot(Petal.Length, Petal.Width, col=Species, data = iris) + geom_hline(yintercept = 1.75) + geom_vline(xintercept = 2.45)

###### we have low data, so only crocc-validation is possible here

## tree
fit <- train(Species ~., data = iris, method="rpart",
             trControl = trainControl("cv", number = 7))
fit
fit$finalModel
fancyRpartPlot(fit$finalModel)
varImp(fit)


## forest
fit2 <- train(Species ~., data = iris, method="rf", ntree=10,
              trControl = trainControl("cv", number = 7))
fit2
varImp(fit2)
```



### loan prediction
Variable Description:  
**Loan_ID**	Unique Loan ID  
**Gender**	Male/ Female  
**Married**	Applicant married (Y/N)  
**Dependents**	Number of dependents  
**Education**	Applicant Education (Graduate/ Under Graduate)  
**Self_Employed**	Self employed (Y/N)  
**ApplicantIncome**	Applicant income  
**CoapplicantIncome**	Coapplicant income  
**LoanAmount**	Loan amount in thousands  
**Loan_Amount_Term**	Term of loan in months  
**Credit_History**	credit history meets guidelines  
**Property_Area**	Urban/ Semi Urban/ Rural  
**Loan_Status**	Loan approved (Y/N)  

```{r loan}
## data
url_train <- "https://raw.githubusercontent.com/yew1eb/DM-Competition-Getting-Started/master/AV-loan-prediction/train.csv"
url_test  <- "https://raw.githubusercontent.com/yew1eb/DM-Competition-Getting-Started/master/AV-loan-prediction/test.csv"

training <- read.csv(url(url_train))
testing <- read.csv(url(url_test))

training <- training[complete.cases(training),]
training <- training[,-1]

## exploratory
plot(training[,-12], col= training$Loan_Status)

###### we have low data + testing data has no output variable for comparasion => only cross-validation

## tree
set.seed(123)
model_fit1 <- train(Loan_Status ~., data = training, method = "rpart",
                    trControl = trainControl("cv", number = 7))
model_fit1
model_fit1$finalModel
fancyRpartPlot(model_fit1$finalModel)
plot(model_fit1)
varImp(model_fit1)
plot(varImp(model_fit1))

predict(model_fit1, testing)


## random forest
model_fit2 <- train(Loan_Status ~., data = training, method = "rf", ntree = 20)
model_fit2
plot(model_fit2)
varImp(model_fit2)
plot(varImp(model_fit2))

predict(model_fit2, testing)


## compare tree predictions vs. rf predictions
confusionMatrix(predict(model_fit1, testing), predict(model_fit2, testing))